---
title: "Assignment_2"
author: "Joslyn Fritz"
date: "5/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(naniar)
library(data.table)
library(DBI)

db <- dbConnect(RSQLite::SQLite(), dbname = "survey.db")

```

```{r}
getwd()

```

**Read in Data and Clean Up**

```{r read in datasets}

#########################################################################################
# Read in files 

parcels_raw <- read_csv("parcels.csv.zip",
                        col_types = cols(.default = col_character())) # read in the parcels data
# parcels data, full zip file

parcels <- parcels_raw %>% 
  transmute(APN, 
         Situs1, 
         Situs2, 
         Acreage = as.numeric(Acreage), 
         UseCode, 
         NonTaxCode, 
         AgPres, 
         LandValue = as.numeric(LandValue), 
         Net_Impr = as.numeric(Net_Impr), 
         Net_AV = as.numeric(Net_AV), 
         M_Address1, 
         M_Address2)

#########################################################################################
# Create a written document of the parcel and usecodes table using "|" as the delimitter.

parcel_written <- write_delim(parcels,
  "Santa_Barbara_County_parcels_2011.csv",
  delim = "|",
  na = "")
parcel_written

```


**Analyze Data Under Specific Circumstances**

**1. What are the 10 most-frequently-occuring land uses (in descending order)?**

```{r 1, most frequenct land uses}



```


**2. How many acres are in agricultural preserves?**

```{r 2, acres in ag preserves}

ag_acre <- parcels %>% 
  filter(!is.na(AgPres)) %>% 
  select(Acreage, AgPres) 
  
ag_acre_sum <- sum(ag_acre$Acreage)
# 547153.3501
# Summed the acres in the agricultural preserve

```


**3. What is the mean net assessed value per acre of the entire county?**

```{r 3, mean net assessed value per acre of county}

value_per_acre <- parcels %>% 
 select(Acreage, Net_AV) 

value_per_acre

# Reorganized dataframe

NetAV_sum<- (sum(value_per_acre$Net_AV))

acre_sum <- sum(value_per_acre$Acreage) 

# Summed up the Net_AV and the Acreage

NAV_per_acre <- NetAV_sum/acre_sum
NAV_per_acre 

```


**4. What is the total net assessed value of all non-taxable parcels?**

```{r 4, total net value of all non-tax parcels}

nontax_parcels <- parcels %>% 
  select(NonTaxCode, Net_AV) %>% 
  filter(!is.na(NonTaxCode))

nontax_parcels_sum <- sum(nontax_parcels$Net_AV)
nontax_parcels_sum
# 1,093,026,091

```


**5. What are the 10 largest property holders, by acreage?**

```{r 5, ten largest property holders, by acreage}

largest_prop_byacre <- parcels %>% 
  select(Acreage, M_Address1, M_Address2) %>% 
  filter(!is.na(M_Address1)) %>% 
  filter(!is.na(M_Address2)) %>% 
  group_by(M_Address1, M_Address2) %>% 
  summarize(sum_acreage = sum(Acreage)) %>% 
  arrange(-sum_acreage) %>% # descending order (-)
  head(10) # shows top ten

# Listing the top 10 largest property holders by acreage

```


**6. What are the 10 largest property holders, by net assessed value?**

```{r 6, ten largest prop holders, by Net_AV}

parcel_prop_NAV <- parcels %>% 
  select(Net_AV, M_Address1, M_Address2) %>% 
  filter(!is.na(M_Address1)) %>% 
  filter(!is.na(M_Address2)) %>% 
  group_by(M_Address1, M_Address2) %>% 
  summarize(sum_NAV = sum(Net_AV)) %>% 
  arrange(-sum_NAV) %>% 
  head(10) 

# Listing the 10 largest property holders by Net Assesed Value in descending order


```



