---
title: "HW1"
author: "Joslyn Fritz"
date: "5/4/2019"
output: html_document
---
**Install Packages**

```{r setup and read in datasets, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

**Read in Data and Clean Up**

```{r read in datasets}

#########################################################################################
# Read in files 

parcels_raw <- read_csv("parcels.csv.zip",
                        col_types = cols(.default = col_character())) # read in the parcels data
# parcels data, full zip file

parcels <- parcels_raw %>% 
  transmute(APN, 
         Situs1, 
         Situs2, 
         Acreage = as.numeric(Acreage), 
         UseCode, 
         NonTaxCode, 
         AgPres, 
         LandValue = as.numeric(LandValue), 
         Net_Impr = as.numeric(Net_Impr), 
         Net_AV = as.numeric(Net_AV), 
         M_Address1, 
         M_Address2)

#########################################################################################
# Create a written document of the parcel and usecodes table using "|" as the delimitter.

parcel_written <- write_delim(parcels,
  "Santa_Barbara_County_parcels_2011.csv",
  delim = "|",
  na = "")
parcel_written

```


**Analyze Data Under Specific Circumstances**

**1. What are the 10 most-frequently-occuring land uses (in descending order)?**

```{r one most frequenct land uses}

usecodes <- read_delim("UseCodes.csv", "|", escape_double = FALSE, trim_ws = TRUE) %>% 
  select("UseCode", "CodeDesc") %>% 
  replace_with_na(replace = list(CodeDesc = c("No Description Available")))
# read in usecodes file

# ,col_types = cols(.default = col_character()), delim = "."

parcel_10 <- parcels %>% 
  left_join(usecodes, by = "UseCode") %>% 
  filter(!is.na(CodeDesc)) %>% 
  group_by(UseCode) %>% 
  count(CodeDesc) %>% 
  arrange(-n) %>% # summarize() will not work here because they are not numeric values.
  head(10) 
parcel_10
# Joined parcels and usecodes



```


**2. How many acres are in agricultural preserves?**

```{r two acres in ag preserves}

ag_acre <- parcels %>% 
  filter(!is.na(AgPres)) %>% 
  select(Acreage, AgPres) 
  
ag_acre_sum <- sum(ag_acre$Acreage)
# 549,563.3511
# Summed the acres in the agricultural preserve

```


**3. What is the mean net assessed value per acre of the entire county?**

```{r three mean net assessed value per acre of county}

value_per_acre <- parcels %>% 
 mutate(Net_AV/Acreage) %>% 
  filter(!is.na(Net_AV/Acreage)) %>% 
  filter(!is.infinite(Net_AV/Acreage))

value_per_acre

mean(value_per_acre$`Net_AV/Acreage`)
# 1,635,938

```


**4. What is the total net assessed value of all non-taxable parcels?**

```{r four total net value of all non-tax parcels}

nontax_parcels <- parcels %>% 
  select(NonTaxCode, Net_AV) %>% 
  filter(!is.na(NonTaxCode))

nontax_parcels_sum <- sum(nontax_parcels$Net_AV)
nontax_parcels_sum
# 1,093,026,091

```


**5. What are the 10 largest property holders, by acreage?**

```{r five ten largest property holders, by acreage}

largest_prop_byacre <- parcels %>% 
  select(Acreage, M_Address1, M_Address2) %>% 
  filter(!is.na(M_Address1)) %>% 
  filter(!is.na(M_Address2)) %>% 
  group_by(M_Address1, M_Address2) %>% 
  summarize(sum_acreage = sum(Acreage)) %>% 
  arrange(-sum_acreage) %>% # descending order (-)
  head(10) # shows top ten
largest_prop_byacre

# Listing the top 10 largest property holders by acreage

```


**6. What are the 10 largest property holders, by net assessed value?**

```{r six ten largest prop holders, by Net_AV}

parcel_prop_NAV <- parcels %>% 
  select(Net_AV, M_Address1, M_Address2) %>% 
  filter(!is.na(M_Address1)) %>%  # !is.na filters everything that is NOT na
  filter(!is.na(M_Address2)) %>% 
  group_by(M_Address1, M_Address2) %>% 
  summarize(sum_NAV = sum(Net_AV)) %>% 
  arrange(-sum_NAV) %>% 
  head(10) 
parcel_prop_NAV

# Listing the 10 largest property holders by Net Assesed Value in descending order


```


