```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
getwd()

```


connect Rmd SQL chunk to SQLite database

```{r}

library(DBI)
db <- dbConnect(RSQLite::SQLite(), dbname = "survey.db")

# Making a connection to SQLite so we can work with it and R at the same time. 
```

## 1. [Selecting Data](https://swcarpentry.github.io/sql-novice-survey/01-select/index.html)

(allows us to work in a bash, because we called bash in the beginning - in the terminal!)
```{bash}
echo '.tables' | sqlite3 survey.db

# Firing up a connect to the survey.db. Started by using "echo", after you fire this up and make the conection it goes away as soon as you're done with it. 
```
(This is us firing it up via SQL, the argument is telling it what connection we're going to use, db. Shows us the contents of a table with similar headings to the bash one. We're asking for a particular column and name.)
```{sql connection=db}
SELECT name AS tables FROM sqlite_master WHERE type = 'table'
```

```{bash}
echo '.schema' | sqlite3 survey.db
```
(SQL gives us info from tables, while the bash gives us info from the terminal)
```{sql connection=db}
SELECT sql AS schema FROM sqlite_master
```

```{sql connection=db}
SELECT family, personal FROM Person;
```
(SQL is not case sensitive)
```{sql connection=db}
SeLeCt FaMiLy, PeRsOnAl FrOm PeRsOn;
```

```{sql connection=db}
SELECT personal, family FROM Person;
```

(If you specificy a column more than once you will get a column more than once! Here it created the new column with 2 or 3 added. This is like joining tables in GIS.)
```{sql connection=db}
SELECT id, id, id FROM Person;
```

(limit 3 will only show us the first 3 things of the table)
(basc syntax = SELECT columns FROM what)
```{sql connection=db}
SELECT * FROM Person LIMIT 3;
```

## 2. [Sorting and Removing Duplicates](https://swcarpentry.github.io/sql-novice-survey/02-sort-dup/index.html)
(Looking at what quantities we want from the Survey table)
```{sql connection=db}
SELECT quant FROM Survey;
```

(We can ask questions, what are the distinct quants we use in this table, and the SQL answers us. It's  very clunky system but does what we need.)
```{sql connection=db}
SELECT DISTINCT quant FROM Survey;
```

(What kinds of surveys were taken at which quantities)
```{sql connection=db}
SELECT DISTINCT taken, quant FROM Survey;
```

(The case doesn't matter for the headers, or the languages, but it does matter for the values!)
```{sql connection=db}
SELECT * FROM Person ORDER BY id;
```

```{sql connection=db}
SELECT * FROM person ORDER BY id DESC;
```

(Know how this is broken down, we called the 3 columns and had them ordered in the specific ways described below.)
```{sql connection=db}
SELECT taken, person, quant FROM Survey ORDER BY taken ASC, person DESC;
```

```{sql connection=db}
SELECT DISTINCT quant, person FROM Survey ORDER BY quant ASC;
```

```{sql connection=db}
select distinct dated from visited; 
```


## 3. [Filtering](https://swcarpentry.github.io/sql-novice-survey/03-filter/index.html)
(This will let us filter things out of rows. You contrain values in rows. We quote the 'DR-1' because it forces it to keep it in a string of values. The where clause will drastically reduce the amount of data you're pulling from the file.)
```{sql connection=db}
SELECT * FROM Visited WHERE site = 'DR-1';
```

```{sql connection=db}
SELECT id FROM Visited WHERE site = 'DR-1';
```

(Both halves of the expression need to be true for it to give an output.)
```{sql connection=db}
SELECT * FROM Visited WHERE site = 'DR-1' AND dated < '1930-01-01';
```

```{sql connection=db}
SELECT * FROM Survey WHERE person = 'lake' OR person = 'roe';
```

```{sql connection=db}
SELECT * FROM Survey WHERE person IN ('lake', 'roe');
```

(The AND binds tighter than the OR, so sal AND person are bound together here. Frew recommends using parenthesis to make it really clear.)
```{sql connection=db}
SELECT * FROM Survey WHERE quant = 'sal' AND person = 'lake' OR person = 'roe';
```

```{sql connection=db}
SELECT * FROM Survey WHERE quant = 'sal' AND (person = 'lake' OR person = 'roe');
```

(% means from here on in, match anything. One of the few times "" are not seen as specifically what you mean. The LIKE operator wants you to interpret the "DR%" as something special.)
```{sql connection=db}
SELECT * FROM Visited WHERE site LIKE 'DR%';
```

```{sql connection=db}
SELECT DISTINCT person, quant FROM Survey WHERE person = 'lake' OR person = 'roe';
```

(This one messed up because we used OR when we should've used AND - the OR grabs everything when we need it to be more precise. When using AND both apply.)
```{sql connection=db}
SELECT * FROM Site WHERE (lat > -48) OR (lat < 48);

```
(Now we get the one site that is within our constraints.)
```{sql connection=db}
SELECT * FROM Site WHERE (lat > -48) AND (lat < 48);

```


## 4. [Calculating New Values](https://swcarpentry.github.io/sql-novice-survey/04-calc/index.html)

(this does not affect the original table, it's not saved in the database. It only does this change for us, right now.)
```{sql connection=db}
SELECT 1.05 * reading FROM Survey WHERE quant = 'rad';
```

(The 2 in the round function rounds it to two places after the decimal.)
```{sql connection=db}
SELECT taken, round(5 * (reading - 32) / 9, 2) FROM Survey WHERE quant = 'temp';
```

(We want the code in caps because it's the example he gives us haha)
```{sql connection=db}
SELECT taken, round(5 * (reading - 32) / 9, 2) AS Celsius FROM Survey WHERE quant = 'temp';
```

(Creates new kinds of columns. This one is the concatination of the family name. Using the AS allows you to write in the new name.)
```{sql connection=db}
SELECT personal || ' ' || family AS full_name FROM Person;
```

(The union select could be switched for OR here but he wants to show us how it works.)
```{sql connection=db}
SELECT * FROM Person WHERE id = 'dyer' UNION SELECT * FROM Person WHERE id = 'roe';
```

```{sql connection=db}
SELECT DISTINCT site FROM Visited;
```

## 5. [Missing Data](https://swcarpentry.github.io/sql-novice-survey/05-null/index.html)

```{sql connection=db}
SELECT * FROM Visited;
```

```{sql connection=db}
SELECT * FROM Visited WHERE dated < '1930-01-01';
```

```{sql connection=db}
SELECT * FROM Visited WHERE dated >= '1930-01-01';
```

```{sql connection=db}
SELECT * FROM Visited WHERE dated = NULL;
```

```{sql connection=db}
SELECT * FROM Visited WHERE dated != NULL;
```
(IS gets us to the NULL we were looking for!)
```{sql connection=db}
SELECT * FROM Visited WHERE dated IS NULL;
```
(Shows everything that is not an N/A)
```{sql connection=db}
SELECT * FROM Visited WHERE dated IS NOT NULL;
```

```{sql connection=db}
SELECT * FROM Visited WHERE dated IN ('1927-02-08', NULL);
```

(Didn't finish this one)
```{sql connection=db}
SELECT * from visited where dated is not null order by asc;
```


## 6. [Aggregation](https://swcarpentry.github.io/sql-novice-survey/06-agg/index.html)

```{sql connection=db}
SELECT dated FROM Visited;
```

```{sql connection=db}
SELECT min(dated) AS earliest FROM Visited;
```

```{sql connection=db}
SELECT max(dated) FROM Visited;
```

```{sql connection=db}
SELECT avg(reading) FROM Survey WHERE quant = 'sal';
```

```{sql connection=db}
SELECT count(reading) FROM Survey WHERE quant = 'sal';
```

```{sql connection=db}
SELECT sum(reading) FROM Survey WHERE quant = 'sal';
```

(Min and max!)
```{sql connection=db}
SELECT min(reading), max(reading) FROM Survey WHERE quant = 'sal' AND reading <= 1.0;
```

(Putting a person here can be misleading. You may want to group by.)
```{sql connection=db}
SELECT person, count(*) FROM Survey WHERE quant = 'sal' AND reading <= 1.0;
```

```{sql connection=db}
SELECT person, max(reading), sum(reading) FROM Survey WHERE quant = 'missing';
```

```{sql connection=db}
SELECT min(dated) FROM Visited;
```

```{sql connection=db}
SELECT min(dated) FROM Visited WHERE dated IS NOT NULL;
```

(Common style to write SQL when they get too long)
```{sql connection=db}
SELECT person, count(reading), round(avg(reading), 2)
FROM  Survey
WHERE quant = 'rad';
```

```{sql connection=db}
SELECT person, count(reading), round(avg(reading), 2)
FROM  Survey
WHERE quant = 'rad'
AND   person = 'dyer';
```

(Group by puts them all into piles, aggregating the rad values, but they're ordered by person)
```{sql connection=db}
SELECT   person, count(reading), round(avg(reading), 2)
FROM     Survey
WHERE    quant = 'rad'
GROUP BY person;
```

```{sql connection=db}
SELECT   person, quant, count(reading), round(avg(reading), 2)
FROM     Survey
GROUP BY person, quant;
```

```{sql connection=db}
SELECT   person, quant, count(reading), round(avg(reading), 2)
FROM     Survey
WHERE    person IS NOT NULL
GROUP BY person, quant
ORDER BY person, quant;
```

(Why did we only get one value back? It collapsed the result set into a single row and subtracted the average from that. How we get this to subtract it from every value?)
```{sql connection=db}
SELECT reading - avg(reading) FROM Survey WHERE quant = 'rad';
```

(This is called a nested Select - it answers our previous question. The () make all of the difference. You could use AS here to create a nicer name for it.)
```{sql connection=db}
SELECT reading - (SELECT avg(reading) FROM Survey WHERE quant = 'rad')
FROM   Survey
WHERE  quant = 'rad'
```

## 7. [Combining Data](https://swcarpentry.github.io/sql-novice-survey/07-join/index.html)

(This gave us every possible join of these tables.)
```{sql connection=db}
SELECT * FROM Site JOIN Visited;
```

(column.table (below) = how we join two tables. Don't ever use . in names of variables in R! Becuase it has a purpose. We made it fit into a table because of the replicating data. Site column is a foreign key that makes the whole join structure work.)
```{sql connection=db}
SELECT * FROM Site JOIN Visited ON Site.name = Visited.site;
```

```{sql connection=db}
SELECT Site.lat, Site.long, Visited.dated
FROM   Site JOIN Visited
ON     Site.name = Visited.site;
```

```{sql connection=db}
SELECT Site.lat, Site.long, Visited.dated, Survey.quant, Survey.reading
FROM   Site JOIN Visited JOIN Survey
ON     Site.name = Visited.site
AND    Visited.id = Survey.taken
AND    Visited.dated IS NOT NULL;
```

```{sql connection=db}
SELECT rowid, * FROM Person;
```

(Looking for south of 49 degrees, and the rest of the constraints. The WHERE clause pulls operators out of the table. To do a join, you need the columns to match. WHERE ones are more simple.)
```{sql connection=db}
SELECT Site.name
FROM   Site JOIN Visited
ON     Site.lat < -49.0
AND    Site.name = Visited.site
AND    Visited.dated >= '1932-01-01';
```


```{r}
query <- "SELECT Site.lat, Site.long FROM Site;"

results <- dbGetQuery(db, query)
print(results)

# the same thing we've been doing the whole time, just now in R and not SQL. 
# Results stores it in the R df so you'll have it forever. 

```

